Verifying Demonstrations

### 2020-03-11 paper/DB-Schnorr-nbit.tex
Edits nbit section
Edits protocol description, l-bit problem section.

\section{Bit-by-bit distance-bounding Schnorr protocol}%

One protocol exists in the literature that provides \acp{PPK} with most of the desired properties described previously, ProProx~\cite{ProProx}\footnote{Note that~\cite{ProProx} uses the abbreviation PoPoK, we prefer \acs{PPK} for shorter notation.}.
ProProx is secure against a malicious verifier, but it provides \iac{PPK} protocol for quadratic residues (\ie protocols of the form \(\PPK[\alpha][a = 
  \alpha^2]\)), while in our context, we need \iac{PPK} protocol for discrete logarithms (\ie \(\PPK[\alpha][a = g^\alpha]\)) due to our choice of anonymous credentials whose desirable properties we list in \cref{ZK-anon-cred}.%, which in turn is due to their \ac{PRF} for deriving identifiers from a long-term ID with unlinkability to the original ID but with proof of validity of this derivation. 
To realize this, we will now introduce such a protocol to enable us to have \ac{DB} anonymous credentials.

\subsection{Reattempting a distance-bounding Schnorr protocol}%
\label{DB-Schnorr-nbit}

In the original \ac{DB} paper by \citet{DistanceBounding}, one of the distance-bounding protocols is based on the Schnorr identification scheme~\cite{Schnorr}.
However, this \ac{DB} Schnorr protocol was shown to be prone to distance hijacking~\cite{DistanceHijacking, TamarinDB}, in addition to not being secure against terrorist fraud.
We now propose another way to turn the Schnorr protocol into a public-key \ac{DB} protocol that is secure against \ac{DBMF}, \ac{DBDF}, \ac{DBDH} and \ac{DBTF}.
It is important in our setting that the protocol is a public-key protocol, both to reach strong privacy properties and also because this property implies protection against an impersonating verifier.

We present the protocol in \cref{SchnorrFigure}.
The (cyclic) group with generator \(g\) and order \(q\) are system parameters.
The private key \(\alpha\) with public key \(A = g^\alpha\) are generated once by the prover in the setup phase.

@@

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSprove\leftrightarrow \DBSverify\) \ac{DB} Schnorr protocol for \(\PK[\alpha][A = g^\alpha]\).
    The protocol should be repeated in full to achieve the desired soundness 
    and distance-bounding errors.
  }%
  \label{SchnorrFigure}
\end{figure*}

During one round, the prover commits to a random nonce: more precisely he chooses \(\rho\rgets \ZZ_q\) uniformly at random, computes \(R\gets g^\rho\) and sends \(R\) to the verifier.
The verifier generates two challenges \(c_0\rgets \bin^k, c_1\rgets \bin^k\) and sends them to the prover.
The prover computes \(s_0\gets \rho - c_0\alpha, s_1\gets \rho - c_1\alpha\).
This step is the main difference to the original Schnorr protocol: the verifier selects two challenges and the prover computes two responses --- but still only \emph{one nonce}, \(\rho\).
This is also different from the original Brands-Chaum protocol, in which the 
prover and verifier jointly construct \emph{one} challenge with \emph{one} 
response.

We let \(|q| = \ceil{\log_2 q}\) denote the length of \(q\) in bits.
The verifier will request all \(|q|\) response-bits from one challenge (say 
\(s_b\)) and only \(0 < l\leq |q|\) from the other (\(s_{1-b}\)).
Then the verifier can authenticate the prover by checking if \(R = 
  g^{s_b}A^{c_b}\).

\subsection{Security analysis}

The prover must know the responses for both challenges to successfully pass the \ac{DB} phase.
The reason for the two challenges but only one random nonce is that knowing the responses to both means learning \(\alpha\).
This allows us to prevent \ac{DBTF}.
Bundling the authentication into the distance-bounding phase (difference from Brands-Chaum) prevents \ac{DBDH}, and both adaptations together aim at \ac{DBMF} resistance. %reducing success of MF according to number of repetitions.
The reason we use secret sharing and not \(s_b\) directly, is to bound the \(l\) bits the verifier must learn of \(s_{1-b}\).
Without secret sharing, the verifier learns \(l\) bits of \(s_{1-b}\), but with secret sharing the verifier learns \(l\) bits of \emph{one share}.
Thus, we still have information theoretic security for \(s_{1-b}\) although we reveal the \(l\) bits.

The intuition behind the protocol security is as follows.
To achieve malicious-verifier zero-knowledge, choose \(k\) logarithmic in the security parameter \(\lambda\) and repeat the protocol \(n\) times, such that 
the knowledge error, \(2^{-kn}\), becomes small enough.
Repeating the protocol is also needed to decrease the success probability of some of adversaries presented below.

The \ac{DB} phase protects against distance fraud.
Once the prover has received \(l+1\) challenges set to \(b\), it knows that the remaining challenges must also be \(b\).
\Iac{DBDF} prover must thus wait for the challenge bit \(b_i\) before responding with \(r_i\) for \(l+1\) challenges.
Thus, the probability of successfully guessing the order of the challenge bits is \(2^{-(l+1)}\) per round or, in total, \(2^{-(l+1)n}\).

The \ac{DB} phase also ensures that \iac{DBMF} adversary will fail the \ac{DB} phase for at least one round.
The verifier will send \(|q| + l < 2|q|\) challenges.
By replying, the prover reveals \(|q|+l\) bits distributed over \(s_0, 
  s_1\).
\emph{We assume} that learning \(\ceil{(|q|+l)/2}\) bits of \(s_0\) and 
\(\floor{(|q|+l)/2}\) bits of \(s_1\) yields no advantage for the adversary.
Now, the adversary has two options.
First, the adversary can buffer \(l\) bits of both \(s_0^{(0)}\concat s_0^{(1)}, s_1^{(0)}\concat s_1^{(1)}\) by requesting them from the prover.
However, at this point the adversary must wait for the challenge bit \(b_{l+1}\) from the verifier and then relay that challenge to the prover to receive the correct \(r_i\) --- this relay will be detected.
The alternative is to guess \(b\) and extract \(s_b\) for this challenge from the prover.
Guessing \(b\) yields \(1/2\) probability that the \ac{DBMF} adversary will guess it correctly.
Since we must execute the protocol \(n\) times, the probability to pass all these rounds is reduced to \(2^{-n}\).
If the adversary requests \(s_b\) in full, the adversary can also guess the 
remaining \(|q|-l\) bits of \(s_{1-b}\).
Based on our assumption above, the probability of guessing these bits 
successfully is \(2^{-(|q|-l)}\).
Thus the success probability of the \ac{DBMF} adversary is \emph{at most} 
\(2^{-n}+2^{-(|q|-l)n}\).

The protocol is \ac{DBTF}-resistant. 
Indeed, if the malicious prover gives both responses to the accomplice, the accomplice can compute his secret key.
The probability of success of \ac{DBTF} is thus reduced to guessing \(b\), \ie \(1/2\) per round or, in total, \(2^{-n}\).

The protocol is also secure against distance hijacking due to the fact that the authenticated bitstring is used during the \ac{DB} phase, not the challenge bitstring as it is the case for the original protocol of Brands-Chaum.
Thus the probability of success is equivalent of a collision for the responses 
for the chosen challenge --- in each round.

Finally, it follows from the original Schnorr protocol that this is still \iac{PK} and \ac{ZK}.
\daniel{I should prove this for the composed protocol, I should check the details with Douglas.}

\subsection{The \(l\)-bit problem}

@@

\begin{definition}[DB-Schnorr Problem]%
  \label{DBSchnorr-problem}
  Let \(g\) be a generator for a group of prime order~\(q\).
  Give \(g, q, R = g^\rho A = g^\alpha\) to the adversary.
  Let the adversary choose \(c\) and additionally give him \(\hat s\) such that 
  \(\hat s\) are the \(l\) \emph{least significant} bits of \(s = \rho + 
  c\alpha \pmod q\).
  The adversary wins if he returns \(\hat s'\) such that \(\hat s + \hat s' = s 
  \pmod q\).
\end{definition}

\begin{lemma}
  Given an algorithm~\(\algA_{\DBS}\) that solves the DB-Schnorr Problem 
  (\cref{DBSchnorr-problem}) with probability \(\epsilon\), we can construct an 
  algorithm \(\algA_{\text{DL}}\) that solves the \ac{DLP} with probability 
  \(\epsilon 2^{-(l+k)}\).
\end{lemma}

>>>>

\section{Bit-by-bit distance-bounding Schnorr protocol}%
\label{DB-Schnorr-nbit}

We will now continue with a more traditional (in \ac{DB} terms) version of the 
protocol, one that uses bit-by-bit communication.

%One protocol exists in the literature that provides \acp{PPK} with most of the desired properties described previously, ProProx~\cite{ProProx}\footnote{Note that~\cite{ProProx} uses the abbreviation PoPoK, we prefer \acs{PPK} for shorter notation.}.
%ProProx is secure against a malicious verifier, but it provides \iac{PPK} protocol for quadratic residues (\ie protocols of the form \(\PPK[\alpha][a = 
%  \alpha^2]\)), while in our context, we need \iac{PPK} protocol for discrete logarithms (\ie \(\PPK[\alpha][a = g^\alpha]\)) due to our choice of anonymous credentials whose desirable properties we list in \cref{ZK-anon-cred}.%, which in turn is due to their \ac{PRF} for deriving identifiers from a long-term ID with unlinkability to the original ID but with proof of validity of this derivation. 
%To realize this, we will now introduce such a protocol to enable us to have \ac{DB} anonymous credentials.

We present the protocol in \cref{DB-Schnorr-nbit-figure}.
We will present the protocol as \iac{DB} version of the Schnorr identification 
scheme (\(\PK[\alpha][A = g^\alpha]\)), this is for the sake of exposition.
It can be generalized to \eg CL04~\cite{CLsignatures}.

The (cyclic) group with generator~\(g\) with prime order~\(q\) is a system 
parameter.
The private key~\(\alpha\) with public key~\(A = g^\alpha\) are generated once 
by the prover in the setup phase.
The verifier has a copy of the public key~\(A\) and wants to verify that the 
private key~\(\alpha\) is within a certain distance-bound.

@@

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSprove\leftrightarrow \DBSverify\) \ac{DB} Schnorr protocol for \(\PK[\alpha][A = g^\alpha]\).
    The protocol should be repeated \(n\) times to achieve the desired 
    soundness and distance-bounding errors.
  }%
  \label{DB-Schnorr-nbit-figure}
\end{figure*}

During one round, the prover commits to a random nonce: more precisely he 
chooses \(
  \rho\rgets \ZZ_q
\) uniformly at random, computes \(
  R\gets g^\rho
\) and sends \(R\) to the verifier.
The verifier generates two challenges \(
  c_0\rgets \bin^k,
  c_1\rgets \bin^k
\) and sends them to the prover.
The prover computes \(
  s_0\gets \rho - c_0\alpha,
  s_1\gets \rho - c_1\alpha.
\)
This step is the main difference to the original Schnorr protocol: the verifier selects two challenges and the prover computes two responses --- but still only \emph{one nonce}, \(\rho\).
This is also different from the original Brands-Chaum protocol, in which the 
prover and verifier jointly construct \emph{one} challenge with \emph{one} 
response.

We let \(|q|\) denote the length of \(q\) in bits, so any element~\(z\in 
\ZZ_q\) can be represented by \(|q|\) bits.
The verifier will request all \(|q|\) response-bits from one challenge (say 
\(s_b\)) and only \(0 < l\leq |q|\) from the other (\(s_{1-b}\)).
Then the verifier can authenticate the prover by checking if \(
  R = g^{s_b}A^{c_b}.
\)

This must be repeated for \(n\) rounds.

\subsection{Security of the \(l\) bits}

Before we proceed to other security properties, we will start with the 
\(l\)-bit problem.
We will now analyse how much damage leaking those \(l\) bits of the second 
response does.

@@

\begin{definition}[DB-Schnorr Problem]%
  \label{DBSP}
  Let \(g\) be a generator for a group of prime order~\(q\).
  Give \(g, q, R = g^\rho A = g^\alpha\) to the adversary.
  Let the adversary choose \(c\) and additionally give him \(\hat s\) such that 
  \(\hat s\) are the \(l\) \emph{least significant} bits of \(s = \rho + 
  c\alpha \pmod q\).
  The adversary wins if he returns \(\hat s'\) such that \(\hat s + \hat s' = s 
  \pmod q\).
\end{definition}

\begin{lemma}\label{DBSPtoDLP}
  Given an algorithm~\(\algA_{\DBS}\) that solves the DB-Schnorr Problem 
  (\cref{DBSP}) with probability \(\epsilon\), we can construct an 
  algorithm \(\algA_{\text{DL}}\) that solves the \ac{DLP} with probability 
  \(\epsilon 2^{-(l+k)}\).
\end{lemma}
